---
title: 'Reproducible Research: Peer Assessment 1'
author: "Jeferson S Monteiro"
date: "11 de abril de 2019"
output: html_document
---

```{r}
echo = TRUE  
options(scipen = 1) 
getwd()
setwd("H:/DS/REPRODUCIBLE_RESEARCH")
```

### Loading and processing the data
```{r}
data <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
data$month <- as.numeric(format(data$date, "%m"))
noNA <- na.omit(data)
rownames(noNA) <- 1:nrow(noNA)
head(noNA)
dim(noNA)
library(ggplot2)
```

### What is mean total number of steps taken per day?

For this moment, ignore the missing values of the dataset.

 > Make a histogram of the total number of steps taken each day
```{r}
ggplot(noNA, aes(date, steps)) + geom_bar(stat = "Identity", colour = "green", fill = "green", width = 0.9) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram: Total Number of Steps Taken Each Day", x = "Date: ", y = "Total number of steps:")
```

 > Mean and median of total number of steps taken per day

Mean total number of steps taken per day:

```{r}
totalSteps <- aggregate(noNA$steps, list(Date = noNA$date), FUN = "sum")$x
mean(totalSteps)
```
Median total number of steps taken per day:
```{r}
median(totalSteps)
```

### What is the average daily activity pattern?

 > The time series plot (i.e. type = "l") of the 5-minute interval (x-axis), averaged across all days (y-axis)

```{r}
avgSteps <- aggregate(noNA$steps, list(interval = as.numeric(as.character(noNA$interval))), FUN = "mean")
names(avgSteps)[2] <- "meanOfSteps"
ggplot(avgSteps, aes(interval, meanOfSteps)) + geom_line(color = "steelblue", size = 0.8) + labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
```
 > The maximum number of steps in the 5-minute interval, on average.

```{r}
avgSteps[avgSteps$meanOfSteps == max(avgSteps$meanOfSteps), ]
```

### Imputing missing values
 > The total number of rows with NAs:

```{r}
sum(is.na(data))
```
 > The strategy for imputing missing data
 
 The strategy was is to use the mean for that 5-minute interval to fill each NA value in the steps column.

 Create a new dataset that is equal to the original, but with the missing data filled in.

```{r}
newData <- data 
for (i in 1:nrow(newData)) {
    if (is.na(newData$steps[i])) {
        newData$steps[i] <- avgSteps[which(newData$interval[i] == avgSteps$interval), ]$meanOfSteps
    }
}
head(newData)
sum(is.na(newData))
```

* Make a histogram of the total number with, mean and median, after imputed missing values:

```{r}
ggplot(newData, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "steelblue",
                                             fill = "steelblue",
                                             width = 0.7) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")
```

* Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Mean total number of steps taken per day:
```{r}
newTotalSteps <- aggregate(newData$steps, 
                           list(Date = newData$date), 
                           FUN = "sum")$x
newMean <- mean(newTotalSteps)
newMean
```
Median total number of steps taken per day:
```{r}
newMedian <- median(newTotalSteps)
newMedian
```
Compare them with the two before imputing missing data:
```{r}
oldMean <- mean(totalSteps)
oldMedian <- median(totalSteps)
newMean - oldMean
newMedian - oldMedian
```

### Are there differences in activity patterns between weekdays and weekends?

 > The new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day was is created.

```{r}
head(newData)
newData$weekdays <- factor(format(newData$date, "%A"))
levels(newData$weekdays)
levels(newData$weekdays) <- list(weekday = c("Monday", "Tuesday",
                                             "Wednesday", 
                                             "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
levels(newData$weekdays)
table(newData$weekdays)
```

 > The panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
library(lattice)

avgSteps <- aggregate(newData$steps, list(interval = as.numeric(as.character(newData$interval)), weekdays = newData$weekdays), FUN = "mean")

print(avgSteps)

names(avgSteps)[3] <- "meanOfSteps"

xyplot(avgSteps$meanOfSteps ~ avgSteps$interval, avgSteps, layout = c(1, 1), type = "1", xlab = "Interval", ylab = "Number of steps")

```
