---
title: "PA1_template"
author: "Jeferson Santos Monteiro"
date: "13 de abril de 2019"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading and processing the data

```{r}

setwd("H:/DS/REPRODUCIBLE_RESEARCH/")
activity<-read.csv("H:/DS/REPRODUCIBLE_RESEARCH/activity.csv")

head(activity)

```

```{r}

totalSteps<-aggregate(steps~date,data=activity,sum,na.rm=TRUE)
```
### What is mean total number of steps taken per day?

For this moment, ignore the missing values of the dataset.


 > Make a histogram of the total number of steps taken each day
 
```{r}
hist(totalSteps$steps)
```
 
 > Mean and median of total number of steps taken per day
 
Mean total number of steps taken per day:

```{r}
mean(totalSteps$steps)
```

Median total number of steps taken per day:

```{r}
median(totalSteps$steps)
```
### What is the average daily activity pattern?

 > The time series plot (i.e. type = "l") of the 5-minute interval (x-axis), averaged across all days (y-axis)

```{r}
stepsInterval<-aggregate(steps~interval,data=activity,mean,na.rm=TRUE)
plot(steps~interval,data=stepsInterval,type="l")
```

 > The maximum number of steps in the 5-minute interval, on average.
 
```{r}
stepsInterval[which.max(stepsInterval$steps),]$interval

```

### Imputing missing values
 > The total number of rows with NAs:

```{r}
sum(is.na(activity$steps))

```
> The strategy for imputing missing data
 
 The strategy was is to use the mean for that 5-minute interval to fill each NA value in the steps column.

 Create a new dataset that is equal to the original, but with the missing data filled in.
 
```{r}
interval2steps<-function(interval){
    stepsInterval[stepsInterval$interval==interval,]$steps
}
```

New dataset.

```{r}
activityFilled<-activity   # Make a new dataset with the original data
count=0           # Count the number of data filled in
for(i in 1:nrow(activityFilled)){
    if(is.na(activityFilled[i,]$steps)){
        activityFilled[i,]$steps<-interval2steps(activityFilled[i,]$interval)
        count=count+1
    }
}
cat("Total ",count, "NA values were filled.\n\r")  
```

* Make a histogram of the total number with, mean and median, after imputed missing values:

```{r}
totalSteps2<-aggregate(steps~date,data=activityFilled,sum)
hist(totalSteps2$steps)
mean(totalSteps2$steps)
median(totalSteps2$steps)
```

> Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

: The **mean** value is the **same** as the value before imputing missing data because we put the mean value for that particular 5-min interval. The median value shows **a little** difference : but it depends on **where the missing values are**.

### Are there differences in activity patterns between weekdays and weekends?

The new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day was is created.


```{r}
activityFilled$day=ifelse(as.POSIXlt(as.Date(activityFilled$date))$wday%%6==0,
                          "weekend","weekday")
# For Sunday and Saturday : weekend, Other days : weekday 
activityFilled$day=factor(activityFilled$day,levels=c("weekday","weekend"))
```


 > The panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).


```{r}
stepsInterval2=aggregate(steps~interval+day,activityFilled,mean)
library(lattice)
xyplot(steps~interval|factor(day),data=stepsInterval2,aspect=1/2,type="l")
```
